---
title: 'Bios 6301: Assignment 3'
author: "Dannielle Gibson"
output:
  pdf_document: default
  html_document: default
---

*Due Tuesday, 27 September, 1:00 PM*

50 points total.

Add your name as `author` to the file's metadata section.

Submit a single knitr file (named `homework3.rmd`) by email to tianyi.sun@vanderbilt.edu.
Place your R code in between the appropriate chunks for each question.
Check your output by using the `Knit HTML` button in RStudio.

$5^{n=day}$ points taken off for each day late.

### Question 1 ###

**15 points**

Write a simulation to calculate the power for the following study
design.  The study has two variables, treatment group and outcome.
There are two treatment groups (0, 1) and they should be assigned
randomly with equal probability.  The outcome should be a random normal
variable with a mean of 60 and standard deviation of 20.  If a patient
is in the treatment group, add 5 to the outcome.  5 is the true
treatment effect.  Create a linear model for the outcome by the
treatment group, and extract the p-value (hint: see assigment1).
Test if the p-value is less than or equal to the alpha level, which
should be set to 0.05.

Repeat this procedure 1000 times. The power is calculated by finding
the percentage of times the p-value is less than or equal to the alpha
level.  Use the `set.seed` command so that the professor can reproduce
your results.

1. Find the power when the sample size is 100 patients. (10 points)
power is 239/1000*100= 23.9%.

```{r}
n.sim =1000

n <- 100
mean <- 60
sd <- 20
p_value = NULL
treatment.groups <-c(0,1) 

set.seed(2)
for (j in 1:n.sim) {
t.s <- sample(treatment.groups, n, replace=TRUE)
outcome<- rnorm(n, mean, sd)
for (i in 1:n) { 
  if (t.s[i]==1){
    outcome[i]<- outcome[i] +5}
  else {outcome[i] <- outcome[i] }  
}
summary(coef(lm(outcome~t.s)))
p_value[j] = coef(summary(lm(outcome~t.s)))[2,4]
}
p_value <0.05 
sum(p_value <0.05)  

```

1. Find the power when the sample size is 1000 patients. (5 points)
power is 46/1000*100= 4.6%
```{r}
n2.sim =1000

n2 <- 1000
mean2 <- 60
sd2 <- 20
p_value2 = NULL
treatment.groups2 <-c(0,1) 

set.seed(2)
for (j2 in 1:n2.sim) {
t.s2 <- sample(treatment.groups2, n2, replace=TRUE)
outcome2<- rnorm(n2, mean2, sd2)
for (i2 in 1:n2) { 
  if (t.s2[i2]==1){
    outcome2[i]<- outcome2[i2] +5}
  else {outcome2[i2] <- outcome2[i2] }  
}
summary(coef(lm(outcome2~t.s2)))
p_value2[j2] = coef(summary(lm(outcome2~t.s2)))[2,4]
}
p_value2 <0.05 
sum(p_value2 <0.05)  

```
### Question 2 ###

**14 points**

Obtain a copy of the [football-values lecture](https://github.com/couthcommander/football-values).
Save the `2021/proj_wr21.csv` file in your working directory.  Read
in the data set and remove the first two columns.

1. Show the correlation matrix of this data set. (4 points)

```{r}
getwd()
proj_wr21<- data.frame(read.csv(file = 'proj_wr21.csv'))         
df<- proj_wr21[-c(1,2)]
df
cor(df, method= c("pearson"))
```

1. Generate a data set with 30 rows that has a similar correlation
structure.  Repeat the procedure 1,000 times and return the mean
correlation matrix. (10 points)

```{r}
library(MASS)
rho<- cor(df)
vcov<- var(df)
means<- colMeans(df)

df.sim <- mvrnorm(20, mu = means, Sigma = vcov)
df.sim = as.data.frame(df.sim)

(rho.sim<- cor(df.sim))   
rho 
keep.1=0
loops=10000

for (i in 1:loops) {
      df.sim = mvrnorm(20, mu = means, Sigma = vcov)
      keep.1=keep.1+cor(df.sim)/loops  }
keep.1-rho
```


### Question 3 ###

**21 points**

Here's some code:

```{r}
nDist <- function(n = 100) {
    df <- 10
    prob <- 1/3
    shape <- 1
    size <- 16
    list(
        beta = rbeta(n, shape1 = 5, shape2 = 45),
        binomial = rbinom(n, size, prob),
        chisquared = rchisq(n, df),
        exponential = rexp(n),
        f = rf(n, df1 = 11, df2 = 17),
        gamma = rgamma(n, shape),
        geometric = rgeom(n, prob),
        hypergeometric = rhyper(n, m = 50, n = 100, k = 8),
        lognormal = rlnorm(n),
        negbinomial = rnbinom(n, size, prob),
        normal = rnorm(n),
        poisson = rpois(n, lambda = 25),
        t = rt(n, df),
        uniform = runif(n),
        weibull = rweibull(n, shape)
    )
}
```

1. What does this do? (3 points)

    ```{r}
    round(sapply(nDist(500), mean), 2)
    ```
    
    ```
      Sapply is  simplifying the output by converting the list of functions into a vector. The mean of the the functions are rounded to two decimal places for a sample size of 500.  
    ```

1. What about this? (3 points)

    ```{r}
    sort(apply(replicate(20, round(sapply(nDist(10000), mean), 2)), 1, sd))
    
    ```
    
    ```
    `
      Sapply is  simplifying the output by converting the list of functions into a vector. The mean of the the functions are rounded to two decimal places for a sample size of 10000. This will then be replicated 20 times whereby creating a 15by20 or 20by15 matrix. Then the apply function will report the standard deviation of the rows of the functions. Lastly the standard deviations will be sorted from smallest to largest. 
    ```

    In the output above, a small value would indicate that `N=10,000` would provide a sufficent sample size as to estimate the mean of the distribution. Let's say that a value *less than 0.02* is "close enough".

1. For each distribution, estimate the sample size required to simulate the distribution's mean. (15 points)
```{r}

nDist <- function(n = 100) {
    df <- 10
    prob <- 1/3
    shape <- 1
    size <- 16
    list(
        beta = rbeta(n, shape1 = 5, shape2 = 45),
        binomial = rbinom(n, size, prob),
        chisquared = rchisq(n, df),
        exponential = rexp(n),
        f = rf(n, df1 = 11, df2 = 17),
        gamma = rgamma(n, shape),
        geometric = rgeom(n, prob),
        hypergeometric = rhyper(n, m = 50, n = 100, k = 8),
        lognormal = rlnorm(n),
        negbinomial = rnbinom(n, size, prob),
        normal = rnorm(n),
        poisson = rpois(n, lambda = 25),
        t = rt(n, df),
        uniform = runif(n),
        weibull = rweibull(n, shape)
    )
}
round(sapply(nDist(500), mean), 2)

```

Don't worry about being exact. It should already be clear that N < 10,000 for many of the distributions. You don't have to show your work. Put your answer to the right of the vertical bars (`|`) below.

distribution|N
---|---
beta|?
binomial|?
chisquared|?
exponential|?
f|?
gamma|?
geometric|?
hypergeometric|?
lognormal|?
negbinomial|?
normal|?
poisson|?
t|?
uniform|?
weibull|?
